---
title: "CCA"
author: "Julia Golebiowska"
date: "2024-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
CCA  - Canonical correspondence analysis on viral data from the Elbe estuary for each fraction and and data type - viral abundance and expression based on individual marker genes. This analysis revels the potential drivers of communities diversity and activity. 



```{r test}
library(vegan)
library(veganUtils)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(tidyverse)
library(ggforce)
library(stringr)
library(RColorBrewer)
library(cluster)
require("knitr")
opts_knit$set(root.dir = "C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/CCA_marker_gene")
setwd("C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/CCA_marker_gene")
```

## Viral diversity 

##1. Read taxonomy data and unify within clusters 


```{r , echo=FALSE}
#load marker genes
geneabund_drep_marker <- readRDS("C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/vibrant_annot_drep_marker_abund.RDS")
#remove repeating rows - I don't need host taxonomy 
#next time generate special object with viral taxonomy and salinity niche classification
geneabund_drep_marker <- geneabund_drep_marker %>% select(contains("GROS")) %>% distinct()
#drep clusters 
drep_clusters <- read.csv("C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/Cdb.csv", header = TRUE, sep = ",")
#taxonomy
drep_clusters[,"genome"] = sub(".fna", "", drep_clusters[,"genome"])
proviruses_taxonomy <- read.csv("C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/Elbe_viruses_distribution/genomad_proviruses_output/Combined_proviruses_annotate/Combined_proviruses_taxonomy.tsv", sep = "\t")
viruses_taxonomy <- read.csv("C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/Elbe_viruses_distribution/genomad_viruses_output/Combined_viruses_taxonomy.tsv", sep = "\t")
#analyse if similar taxonomy for each cluster
drep_clusters <- left_join(drep_clusters, viruses_taxonomy, by=c("genome"="seq_name"))
drep_clusters %>% group_by(primary_cluster) %>% summarise(count_genomes= n_distinct(genome), count_lineages = n_distinct(lineage))  %>%  filter(count_lineages > 1) 
#only 46 have more then 1 lineage - even when more lineages they agree at some point 
#lets unify those 
#keep lineages which dominates by more then 50%
#if less then 50% go to upper taxa level
#continue until Viruses - as we analyse only viruses we will put -"other" label to all analysis
#for this purpose separate into columns
drep_clusters <- drep_clusters %>% 
  separate(lineage, c("level_1", "level_2", "level_3", "level_4", "level_5",  "level_6", "level_7"), ";")
drep_clusters[drep_clusters == ""] <- NA
columns_to_keep = c("primary_cluster", "genome", "level_1", "level_2", "level_3", "level_4", "level_5",  "level_6", "level_7")
#primary_cluster, genome, level_1, level_2, level_3, level_4, level_5, level_6, level_7
drep_clusters_fin_taxonomy <- drep_clusters %>%  add_count(primary_cluster, level_7,  name = "level_7_percentage", wt = !is.na(level_7))  %>% ungroup() %>%
  add_count(primary_cluster, level_6, wt = !is.na(level_6), name = "level_6_percentage") %>% ungroup() %>%
  add_count(primary_cluster, level_5, wt = !is.na(level_5), name = "level_5_percentage") %>% ungroup() %>%
  add_count(primary_cluster, level_4, wt = !is.na(level_4), name = "level_4_percentage") %>% ungroup() %>%
  add_count(primary_cluster, level_3, wt = !is.na(level_3), name = "level_3_percentage") %>% ungroup() %>% 
  add_count(primary_cluster, level_2, wt = !is.na(level_2), name = "level_2_percentage") %>% ungroup() %>%
  add_count(primary_cluster, level_1, wt = !is.na(level_1), name = "level_1_percentage") %>% mutate(level_2_percentage = level_2_percentage/level_1_percentage, level_3_percentage = level_3_percentage/level_1_percentage, level_4_percentage = level_4_percentage/level_1_percentage, level_5_percentage = level_5_percentage/level_1_percentage, level_6_percentage = level_6_percentage/level_1_percentage, level_7_percentage = level_7_percentage/level_1_percentage, level_1_percentage = level_1_percentage/level_1_percentage)  
#unify taxon according to pseudocode above
drep_clusters_fin_taxonomy_unified <- drep_clusters_fin_taxonomy %>% group_by(primary_cluster) %>%
  filter(if(any(level_7_percentage  > 0.5,  na.rm = T)) level_7_percentage > 0.5 else TRUE ) %>% ungroup() %>% group_by(primary_cluster) %>% 
  filter(if(any(level_6_percentage  > 0.5,  na.rm = T)) level_6_percentage > 0.5 else TRUE ) %>% ungroup() %>% group_by(primary_cluster) %>%
  filter(if(any(level_5_percentage  > 0.5,  na.rm = T)) level_5_percentage > 0.5 else TRUE ) %>% ungroup() %>% group_by(primary_cluster) %>%
  filter(if(any(level_4_percentage  > 0.5,  na.rm = T)) level_4_percentage > 0.5 else TRUE ) %>% ungroup() %>% group_by(primary_cluster) %>%
  filter(if(any(level_3_percentage  > 0.5,  na.rm = T)) level_3_percentage > 0.5 else TRUE ) %>% ungroup() %>% group_by(primary_cluster) %>%
  filter(if(any(level_2_percentage  > 0.5,  na.rm = T)) level_2_percentage > 0.5 else TRUE ) %>% ungroup() %>% group_by(primary_cluster) %>% 
  filter(if(any(level_1_percentage  > 0.5,  na.rm = T)) level_1_percentage > 0.5 else TRUE ) %>% ungroup() %>% select(-c("genome","secondary_cluster","threshold", "cluster_method", "comparison_algorithm", "n_genes_with_taxonomy", "agreement", "taxid")) %>% distinct()
  
drep_clusters_fin_taxonomy_unified <- drep_clusters_fin_taxonomy_unified %>% rowwise() %>% mutate(lineage=case_when(level_7_percentage > 0.5 ~ str_c(c(level_1, level_2, level_3, level_4, level_5, level_6, level_7), collapse=";"), level_6_percentage > 0.5 ~ str_c(c(level_1, level_2, level_3, level_4, level_5, level_6), collapse=";"), level_5_percentage > 0.5 ~ str_c(c(level_1, level_2, level_3, level_4, level_5), collapse=";"), level_4_percentage > 0.5 ~ str_c(c(level_1, level_2, level_3, level_4), collapse=";"),  level_3_percentage > 0.5 ~ str_c(c(level_1, level_2, level_3), collapse=";"), level_2_percentage > 0.5 ~ str_c(c(level_1, level_2), collapse=";"),  level_1_percentage > 0.5  ~ level_1))

#unite lineage and remove repeating rows
drep_clusters_fin_taxonomy_unified = drep_clusters_fin_taxonomy_unified %>% 
  separate("lineage", c("level_1", "level_2", "level_3", "level_4", "level_5",  "level_6", "level_7"), sep = ";", remove = FALSE) %>% select(c("primary_cluster", "level_1", "level_2", "level_3", "level_4", "level_5", "level_6", "level_7", "lineage")) %>% distinct()
#set factors levels of salinity

drep_clusters_fin_taxonomy_unified %>% group_by(primary_cluster) %>% summarise(n=n()) %>% arrange(desc(n))  
#good
#add marker genes counts
geneabund_drep_marker_taxonomy <- left_join(geneabund_drep_marker, drep_clusters_fin_taxonomy_unified, by = "primary_cluster")
#make a taxonomy boxplot
#read metadata 
path = "C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/metadata/"
metadata <- read.csv(paste0(path,"/PhysicochemicalParameters_mod3.csv"), header=TRUE, sep=",", row.names=1) 
metadata$Station <- factor(metadata$Station, levels = c("BunthausSpitze", "Muhlenberger Loch","Twielenfleth", "Schwarztonnensand", "Brunsbuttel","Meedem Grund"))

#sum by lineage and add metadata - but first make it long format 
geneabund_drep_marker_taxonomy_long <- geneabund_drep_marker_taxonomy %>% pivot_longer(cols = starts_with("GROS"), names_to = "sampleid", values_to = "counts", values_drop_na = 0) %>% group_by(lineage, sampleid, level_5, primary_cluster) %>% summarise(counts=sum(counts)) %>% filter(counts>0) %>% left_join(metadata)
# %>% group_by(sampleid, level_5) %>% summarise(counts=sum(counts), level_5, primary_cluster) 
# %>% ungroup() %>% group_by(lineage, sampleid, level_5, primary_cluster) %>% summarise(counts=sum(counts)) 
#add poly- meso- oligo - haline labels
#little change of oligohaline definition till 5.5 - according to k-medioids analysis
# res <- pam(metadata$Salinity_PSU, 4)
# res$clustering
#  max(metadata$Salinity_PSU[res$clustering==1])
# [1] 1.39 
# min(metadata$Salinity_PSU[res$clustering==2])
# [1] 12.043
# min(metadata$Salinity_PSU[res$clustering==3])
# [1] 5.02
# min(metadata$Salinity_PSU[res$clustering==4])
# [1] 19.96339
#why not on meta data - less calculation
geneabund_drep_marker_taxonomy_long <- geneabund_drep_marker_taxonomy_long %>% mutate(Salinity_level = case_when(Salinity_PSU <= .5  ~ 'freshwater',  Salinity_PSU > .5 & Salinity_PSU <= 5.5  ~ 'oligohaline', Salinity_PSU > 5.5 & Salinity_PSU <= 18  ~ 'mesohaline', Salinity_PSU > 18  ~ 'polyhaline'))
#remove stations  - counts without metadata 
#test if we have one line per cluster
#think about finding metadata for them 
geneabund_drep_marker_taxonomy_long <- geneabund_drep_marker_taxonomy_long %>% filter(!is.na(Station))
#change NA to unknown label for plotting
geneabund_drep_marker_taxonomy_long <- geneabund_drep_marker_taxonomy_long %>% mutate(level_5 = replace_na(level_5, "aunknown")) %>% distinct()
saveRDS(geneabund_drep_marker_taxonomy_long, file ="geneabund_drep_marker_taxonomy_long.RDS")
#box plot
#simplified lineages for plotting
#lets make NA transparent 
geneabund_drep_marker_taxonomy_long$level_5 <- as.factor(geneabund_drep_marker_taxonomy_long$level_5)
levels(geneabund_drep_marker_taxonomy_long$level_5) <- c("aunknown", "Caudoviricetes", "Duplopiviricetes", "Megaviricetes",  "Maveriviricetes",  "Polintoviricetes")

taxa_seasons <- ggplot(geneabund_drep_marker_taxonomy_long) + geom_bar(position ="fill", stat = "identity", aes(x=Station, y=counts, fill = level_5)) + scale_fill_manual(name ="Class", values=c( "NA", brewer.pal(6, "Set2")), na.value=NA, labels = c("", "Caudoviricetes", "Duplopiviricetes", "Megaviricetes",  "Maveriviricetes",  "Polintoviricetes"), na.translate = FALSE) +
  scale_y_continuous(labels=scales::percent_format()) + facet_wrap(~data_type *Sample_date) + ggtitle("Relative abundace across different seasons") + ylab("Relative abundance") + theme(axis.title.y = element_text(angle=90, size=10), axis.title.x = element_blank(), axis.text.x = element_text(angle=90, hjust=1, size=10), legend.key = element_rect(fill = "transparent"))
tiff("Relative_abundace_taxonomy_across_different_seasons.tiff", unit="px", width = 1000, height = 1000)
taxa_seasons
dev.off()
 tiff("Relative_abundace_taxonomy_across_different_fraction.tiff", unit="px", width = 1000, height = 1000)
ggplot(geneabund_drep_marker_taxonomy_long) + geom_bar(position ="fill", stat = "identity", aes(x=Station, y=counts, fill = level_5)) + scale_y_continuous(labels=scales::percent_format()) + facet_wrap(~data_type *Sample_type) + ggtitle("Relative abundace across different fractions")+ ylab("Relative abundance") + theme(axis.title.y = element_text(angle=90, size=12), axis.title.x = element_blank(), axis.text.x = element_text(angle=90, hjust=1, size=12)) + scale_fill_brewer(palette = "Set2") + guides(fill=guide_legend(title="Viral class")) 
dev.off()
#investigate low abundance viruses pointoviricetes

#ggplot(data.frame(geneabund_drep_marker_taxonomy_long[which(geneabund_drep_marker_taxonomy_long$level_5=="Polintoviricetes"),c(5, 7,9,11,12,13)]), aes(x=Station, y=counts)) + geom_bar(stat='identity' ) + facet_grid(vars(Sample_date))

viruses_taxonomy_salinity <- geneabund_drep_marker_taxonomy_long %>% filter(counts!=0 ) %>% ggplot() + geom_bar(position ="fill", stat = "identity",  na.rm = TRUE, aes(x=AccessionNumber_TBDSven, y=counts, fill = level_5)) + scale_y_continuous(labels=scales::percent_format()) + facet_wrap(~data_type *Sample_type, scales = "free") + #facet_grid(rows=vars(data_type), cols=vars(Salinity_level), scales= "free") + 
  ggtitle("Relative abundace across different fractions") + ylab("Relative abundace") + theme(axis.title.y = element_text(angle=90, size=20), axis.title.x=element_blank(), axis.text.x=element_blank()) + guides(fill=guide_legend(title="Viral class")) + scale_x_discrete(labels = NULL, breaks = NULL) + scale_fill_brewer(palette = "Set2") #+ theme_minimal()

tiff("Relative_abundace_taxonomy_across_different_salinity.tiff", unit="px", width = 1000, height = 1000)
viruses_taxonomy_salinity
dev.off()
```
##2. Characterise clusters by salinity preferences and fractions


```{r}
drep_clusters_fin_taxonomy <- drep_clusters_fin_taxonomy %>% mutate(AccessionNumber_TBDSven = str_extract(genome, "SAMEA[0-9]+"))
drep_clusters_fin_taxonomy_description <- merge(drep_clusters_fin_taxonomy, distinct(geneabund_drep_marker_taxonomy_long[, c("AccessionNumber_TBDSven", "Sample_type", "Salinity_level")]), all.x = T, all.y = F, by="AccessionNumber_TBDSven") 

find_niche <-function(x, threshold){
  #calculate rel frequencies
  tab = prop.table(table(x))
  #check id=f there are values >= 0.8
  #then take this value
  if(any(tab>=threshold)){
    names(tab[which(tab>=threshold)])
  }
  #else concatenate values which are higher the 0.2
  else {
    paste0(names(tab[which(tab>= 0.2)]), collapse = ";")
  }}
#statistics for each cluster
drep_clusters_fin_taxonomy_description <- 
  drep_clusters_fin_taxonomy_description %>% 
  add_count(primary_cluster,  name = "count_genomes", wt = !is.na(genome)) %>% ungroup() %>% 
  add_count(primary_cluster, Salinity_level,  name = "percentage_salinity", wt = !is.na(Salinity_level)) %>% ungroup() %>%
  mutate(percentage_salinity = (percentage_salinity/count_genomes)*100) %>% ungroup() %>% 
  add_count(primary_cluster, Sample_type,  name = "percentage_fraction", wt = !is.na(Sample_type)) %>% ungroup()  %>% 
  mutate(percentage_fraction = (percentage_fraction/count_genomes)*100) %>% ungroup() %>% group_by(primary_cluster) %>% mutate(Salinity_niche= find_niche(Salinity_level, 0.8), Fraction_niche= find_niche(Sample_type, 0.8)) %>% ungroup()
#add data to dataframe with primary clusters only to have all calculated taxonomy, 
drep_clusters_fin_taxonomy_unified <- merge(drep_clusters_fin_taxonomy_unified, distinct(drep_clusters_fin_taxonomy_description[, c("count_genomes", "Salinity_niche", "Fraction_niche", "primary_cluster")]), by="primary_cluster", all.x = T, all.y = F,) 
#ok with prepared table lets do the basic statistics for cluster which are bigger then 5 
drep_clusters_fin_taxonomy_unified$Salinity_niche <- factor(drep_clusters_fin_taxonomy_unified$Salinity_niche, levels = c("freshwater", "freshwater;oligohaline", "oligohaline", "freshwater;mesohaline", "mesohaline;oligohaline", "mesohaline", "mesohaline;polyhaline", "polyhaline", "oligohaline;polyhaline", "mesohaline;oligohaline;polyhaline", ""))
#save as RDS
saveRDS(drep_clusters_fin_taxonomy_unified, file="drep_clusters_fin_taxonomy_unified.rds")
#check correlation between salinity and fraction niche  
corr_spearman <- cor(x=rank(drep_clusters_fin_taxonomy_unified$Salinity_niche), y=rank(drep_clusters_fin_taxonomy_unified$Fraction_niche), method = 'spearman')
#0.2 weak positive corelation
#for statistic choose clusters size of more then 4
drep_clusters_fin_taxonomy_unified %>% filter(count_genomes>4) %>% summary() 
drep_clusters_fin_taxonomy_unified_stat <- drep_clusters_fin_taxonomy_unified %>% filter(count_genomes>4)
#do venn diagrams
library(VennDiagram)
 
# Generate 3 sets of 200 words
#debug
A_freshwater <- drep_clusters_fin_taxonomy_unified_stat$primary_cluster[which(grepl("fresh", drep_clusters_fin_taxonomy_unified_stat$Salinity_niche))]
B_oligohaline <- drep_clusters_fin_taxonomy_unified_stat$primary_cluster[which(grepl("oligo", drep_clusters_fin_taxonomy_unified_stat$Salinity_niche))]
C_mesohaline <- drep_clusters_fin_taxonomy_unified_stat$primary_cluster[which(grepl("meso", drep_clusters_fin_taxonomy_unified_stat$Salinity_niche))]
D_polyhaline <- drep_clusters_fin_taxonomy_unified_stat$primary_cluster[which(grepl("poly", drep_clusters_fin_taxonomy_unified_stat$Salinity_niche))]

library(ggvenn)
colors <- brewer.pal(n = 4, name = "Blues") 
# Chart
ggvenn(list(Freshwater=A_freshwater, Oligohaline=B_oligohaline, Mesohaline=C_mesohaline, Polyhaline=D_polyhaline), fill_color = colors)

#most viral clusters have specificity towards salinity. The least in polyhaline only but these samples has the lowest frequencies. The intersections with mesohaline for both polyhaline and oligohaline has also high values as - mesohaline water are neighbouring for both of these conditions. Interesting there is no intersection for polyhaline and oligohaline water and very few for all three categories what strongly suggest separate viral saline niches. 
###for fraction
# Generate 3 sets of 200 words
#debug
Free_living <- drep_clusters_fin_taxonomy_unified_stat$primary_cluster[which(grepl("Free", drep_clusters_fin_taxonomy_unified_stat$Fraction_niche))]
Light_fraction <- drep_clusters_fin_taxonomy_unified_stat$primary_cluster[which(grepl("Light", drep_clusters_fin_taxonomy_unified_stat$Fraction_niche))]
Heavy_fraction <- drep_clusters_fin_taxonomy_unified_stat$primary_cluster[which(grepl("Heavy", drep_clusters_fin_taxonomy_unified_stat$Fraction_niche))]
 
# Chart
venn.diagram(
  x = list(Free_living, Light_fraction, Heavy_fraction),
  category.names = c("Free-living" , "Light fraction" , "Heavyfraction"),
  filename = 'fraction_venn_diagramm.png',
  output=TRUE
)
# Chart
venn.diagram(
  x = list(Free_living, Light_fraction, Heavy_fraction),
  category.names = c("Free-living" , "Light fraction" , "Heavyfraction"),
  filename = 'fraction_venn_diagramm.tiff',
  output=TRUE
)
tiff("C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/CCA_marker_gene/fraction_venn_diagramm.tiff")
ggvenn(list(`Free-living`=Free_living, `Light fraction`=Light_fraction, `Heavy fraction`=Heavy_fraction))
#In contrary for fractions there is no such specificity. Most clusters are present in all fraction or as a second category in particulate fractions
dev.off()
```
```{r}
#add niche description
#and do plots
geneabund_drep_marker_taxonomy_long <- geneabund_drep_marker_taxonomy_long %>% merge(drep_clusters_fin_taxonomy_description[,c("primary_cluster", "Salinity_niche", "Fraction_niche")], by="primary_cluster")
#take clusters 
# good_clusters <- unique(drep_clusters_fin_taxonomy_description$primary_cluster)
#retrieve clusters which has NA salinity level - means there are assembled from cruise without background parameters
na_clusters <- unique(geneabund_drep_marker_taxonomy_long$primary_cluster[which("" == geneabund_drep_marker_taxonomy_long$Salinity_niche)])
geneabund_drep_marker_taxonomy_long <- geneabund_drep_marker_taxonomy_long[which(!geneabund_drep_marker_taxonomy_long$primary_cluster %in%  na_clusters),]
geneabund_drep_marker_taxonomy_long$level_5 <- as.factor(geneabund_drep_marker_taxonomy_long$level_5)
levels(geneabund_drep_marker_taxonomy_long$level_5) <- c("unknown", "Caudoviricetes", "Duplopiviricetes", "Megaviricetes",  "Maveriviricetes",  "Polintoviricetes")
geneabund_drep_marker_taxonomy_long$Salinity_niche <- factor(geneabund_drep_marker_taxonomy_long$Salinity_niche, levels = c("freshwater", "freshwater;oligohaline", "oligohaline", "freshwater;mesohaline", "mesohaline;oligohaline", "mesohaline", "mesohaline;polyhaline", "polyhaline", "oligohaline;polyhaline", "mesohaline;oligohaline;polyhaline", ""))
Salinity_niche_taxa_plot <- na.omit(geneabund_drep_marker_taxonomy_long[, c("sampleid", "counts", "level_5", "Salinity_level", "data_type", "Salinity_niche")]) %>% filter(Salinity_niche %in% c("freshwater",  "oligohaline", "mesohaline", "polyhaline")) %>% ggplot() + geom_bar(position ="fill", stat = "identity", aes(x=sampleid, y=counts, fill = level_5)) + facet_wrap( ~data_type *Salinity_niche, scales= "free_x", nrow = 2) + scale_y_continuous(labels=scales::percent_format()) + scale_x_discrete(name ="Sample", labels = NULL, breaks = NULL) + ggtitle("Relative abundace across different salinity niches") + ylab("Relative abundance") + theme(legend.key = element_rect(fill = "transparent")) + guides(fill=guide_legend(title="Class"))  + scale_fill_manual(name ="Class", values=c( "NA", brewer.pal(6, "Set2")), na.value=NA, labels = c("", "Caudoviricetes", "Duplopiviricetes", "Megaviricetes",  "Maveriviricetes",  "Polintoviricetes"), na.translate = FALSE)
tiff("Salinity_niche_taxa_plot_1.tiff", unit="px", width = 1000, height = 1000)
Salinity_niche_taxa_plot  + theme(text = element_text(size=20))
dev.off()
#facet_wrap(~data_type *Salinity_niche, scales="free")
Fraction_niche_taxa_plot <- na.omit(geneabund_drep_marker_taxonomy_long[, c("sampleid", "Associatednumber", "counts", "level_5", "data_type", "Fraction_niche")]) %>% filter(counts>0) %>% ggplot() + geom_bar(position ="fill", stat = "identity", aes(x=Associatednumber, y=counts, fill = level_5)) + scale_x_discrete(labels = NULL, breaks = NULL) +
  scale_y_continuous(labels=scales::percent_format())  + ggtitle("Relative abundace across different fraction niches") + ylab("Relative abundace") + theme(axis.title.y = element_text(angle=90, size=12), legend.key = element_rect(fill = "transparent")) + guides(fill=guide_legend(title="Class")) + facet_wrap(~data_type *Fraction_niche, scales= "free_x")
tiff("Fraction_niche_taxa_plot.tiff", unit="px", width = 1000, height = 1000)
Fraction_niche_taxa_plot
dev.off()
```
## 3.Read metadata for each fraction and data type
```{r}

#no as much differences in taxonomy as expected - taking into account PCA of genomic potential viral diversity -
#but there are things to corre -NA  from metadata and different colours for level 3/4 
#FL
viral_expression_FL = readRDS( file ="C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/mantel_test/viral_expression_FL.RDS")
metadata_METAG_FL = readRDS(file ="C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/mantel_test/metadata_METAG_FL.RDS")
metadata_METAT_FL = readRDS(file ="C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/mantel_test/metadata_METAT_FL.RDS")

colnames_FL <- colnames(geneabund_drep_marker)[which(grepl(paste(metadata_METAG_FL$AccessionNumber_TBDSven, collapse="|"), colnames(geneabund_drep_marker)))]
#lets work with data frame 
#keep only clusters which were assembled from samples with background parameters 
geneabund_drep_marker <- as.data.frame(geneabund_drep_marker[which(!geneabund_drep_marker$primary_cluster %in% na_clusters), ])
#set rownames as primary cluster to track particular viral cluster in cca plots
rownames(geneabund_drep_marker) <- geneabund_drep_marker$primary_cluster
colnames_METAG_FL <- colnames_FL[grep("METAG", colnames_FL)]
#keep rownames
viral_diversity_FL <- geneabund_drep_marker[,colnames_METAG_FL]
colnames_METAT_FL <- colnames_FL[grep("METAT", colnames_FL)]
viral_activity_FL <- geneabund_drep_marker[,colnames_METAT_FL] 
#LF

metadata_METAG_LF = readRDS(file ="C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/mantel_test/metadata_METAG_LF.RDS")
metadata_METAT_LF = readRDS(file ="C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/mantel_test/metadata_METAT_LF.RDS")
colnames_LF <- colnames(geneabund_drep_marker)[which(grepl(paste(metadata_METAG_LF$AccessionNumber_TBDSven, collapse="|"), colnames(geneabund_drep_marker)))]
colnames_METAG_LF <- colnames_LF[which(grepl("METAG", colnames_LF))]

#remove na frommcolnames_METAG_LF
viral_diversity_LF <- geneabund_drep_marker[,colnames_METAG_LF] 
colnames_METAT_LF <- colnames_LF[which(grepl("METAT", colnames_LF))]
viral_activity_LF <- geneabund_drep_marker[,colnames_METAT_LF] 
#HF
# viral_expression_HF = readRDS( file ="C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/viral_expression_HF.RDS")
metadata_METAG_HF = readRDS(file ="C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/mantel_test/metadata_METAG_HF.RDS")
metadata_METAT_HF = readRDS(file ="C:/Users/jgolebiowska/Documents/IGB_phd/BICEST/virus/viral_abundances/mantel_test/metadata_METAT_HF.RDS")
colnames_HF <- colnames(geneabund_drep_marker)[which(grepl(paste(metadata_METAG_HF$AccessionNumber_TBDSven, collapse="|"), colnames(geneabund_drep_marker)))]
colnames_METAG_HF <- colnames_HF[which(grepl("METAG", colnames_HF))]

viral_diversity_HF <- geneabund_drep_marker[,colnames_METAG_HF] 
colnames_METAT_HF <- colnames_HF[which(grepl("METAT", colnames_HF))]
viral_activity_HF <- geneabund_drep_marker[,colnames_METAT_HF] 
#vector of numeric parameters columns from metadata
# env_numeric <- colnames(metadata_METAG_LF)[9:40]
# env_numeric <- c("Sat_O2_Perc", "Temperature_TBDHereon", "Salinity_PSU", "Turbidity_NTU", "pH", "O2_uM", "SPM_mgperL", "DOC_mg.L", "TN_mg.L", "DIC_mg.L", "Silicate_mg.L", "Ammonium_mg.L", "Nitrite_mg.L", "Nitrate_mg.L", "SRP_mgperL", "TotalDissolvedPhosphate_mg.L", "Phosphate_uM", "POC_mgperL", "PTC_mgperL", "PTN_mgperL", "PTH_mgperL", "TEP_um2perL", "CSP_um2perL", "ParticleArea_um2perL", "dCH4_uM", "dCO2_uM", "Chlorophyll_Fluoresence", "Chla_ug.l", "bacteria_counts")
env_numeric <- c("Sat_O2_Perc", "Temperature_TBDHereon", "Salinity_PSU", "Turbidity_NTU", "pH", "O2_uM", "SPM_mgperL", "DOC_mg.L", "TN_mg.L", "DIC_mg.L", "Silicate_mg.L", "Ammonium_mg.L", "Nitrite_mg.L", "Nitrate_mg.L", "SRP_mgperL", "TotalDissolvedPhosphate_mg.L", "Phosphate_uM", "POC_mgperL", "PTC_mgperL", "PTN_mgperL", "PTH_mgperL", "TEP_um2perL", "CSP_um2perL", "ParticleArea_um2perL", "dCH4_uM", "dCO2_uM", "Chlorophyll_Fluoresence",  "Chla_ug/l","bacteria_counts")
#for plotting we will change labels
# env_numeric_sim <- c("O2%", "Temp", "Salinity", "Turbidity", "pH", "O2", "SPM", "DOC", "TN", "DIC", "Si", "NH4", "NO2", "NO3", "SRP", "TDP", "PO43-", "POC", "PTC", "PTN", "PTH", "TEP", "CSP", "ParticleArea", "CH4", "CO2", "Chla_Fluoresence", "Chla_ug.l", "bacteria_counts") 
#reformat metadata - sample names into rownames  
#take only numeric parameters to data frame 
metadata_METAG_FL <- metadata_METAG_FL[,env_numeric]
metadata_METAG_FL <- data.frame(apply(metadata_METAG_FL,2, as.numeric))
metadata_METAT_FL <- metadata_METAT_FL[,env_numeric]
metadata_METAT_FL <- data.frame(apply(metadata_METAT_FL,2, as.numeric))

metadata_METAG_LF <- metadata_METAG_LF[,env_numeric]
metadata_METAG_LF <- data.frame(apply(metadata_METAG_LF,2, as.numeric))
metadata_METAT_LF <- metadata_METAT_LF[,env_numeric]
metadata_METAT_LF <- data.frame(apply(metadata_METAT_LF,2, as.numeric))
metadata_METAG_HF <- metadata_METAG_HF[,env_numeric]
metadata_METAG_HF <- data.frame(apply(metadata_METAG_HF,2, as.numeric))
metadata_METAT_HF <- metadata_METAT_HF[,env_numeric]
metadata_METAT_HF <- data.frame(apply(metadata_METAT_HF,2, as.numeric))


#remove column with only NA values
#apply(metadata_METAT_FL,2, FUN=function(x) sum(!is.na(x))/length(x))
not_any_na <- function(x) all(!is.na(x))
#imput pH using O2 saturation and CO2
fit_pH <- lm(pH ~ Sat_O2_Perc + dCO2_uM , metadata_METAG_FL)
#imput SRP
# fit_SRP <- lm(SRP_mgperL ~ Sat_O2_Perc + TN_mg.L + Chlorophyll_Fluoresence, metadata_METAG_FL)
metadata_METAG_FL <- metadata_METAG_FL %>% 
  rowwise() %>% 
  mutate(pH = ifelse(is.na(pH), predict(fit_pH, .cols = c(Sat_O2_Perc, dCO2_uM)), pH)) %>% 
  ungroup()
#remove incomplete columns
METAG_FL_param_keep <- apply(metadata_METAG_FL, 2, not_any_na)
metadata_METAG_FL <- data.frame(metadata_METAG_FL[,METAG_FL_param_keep])

#METAT FL
METAT_FL_param_keep <- apply(metadata_METAT_FL, 2, not_any_na)
metadata_METAT_FL <- data.frame(metadata_METAT_FL[,METAT_FL_param_keep])
#METAG LF
#to impute
#Silicate_mg.L - Temperature_TBDHereon     Turbidity_NTU                O2_uM    DIC_mg.L    Nitrate_mg.L         
#POC_mgperL                   PTC_mgperL                   PTN_mgperL     - Turbidity_NTU                                     
#SRP_mgperL - as above Sat_O2_Perc + TN_mg.L + Chlorophyll_Fluoresence
#imput SRP
fit_SRP <- lm(SRP_mgperL ~ Sat_O2_Perc + TN_mg.L + Chlorophyll_Fluoresence, metadata_METAG_LF)
fit_POC <- lm(POC_mgperL  ~  Turbidity_NTU, metadata_METAG_LF)
fit_PTC <- lm(PTC_mgperL  ~  Turbidity_NTU, metadata_METAG_LF)
fit_PTN <- lm(PTN_mgperL  ~  Turbidity_NTU, metadata_METAG_LF)
fit_Silicate <- lm(Silicate_mg.L ~  Temperature_TBDHereon  + Turbidity_NTU + O2_uM + DIC_mg.L +  Nitrate_mg.L, metadata_METAG_LF)

metadata_METAG_LF <- metadata_METAG_LF %>% 
  rowwise() %>% 
  mutate(POC_mgperL = ifelse(is.na(POC_mgperL), predict(fit_POC, .cols = c(Turbidity_NTU)), POC_mgperL),
         PTC_mgperL = ifelse(is.na(PTC_mgperL), predict(fit_PTC, .cols = c(Turbidity_NTU)), PTC_mgperL),
         PTN_mgperL = ifelse(is.na(PTN_mgperL), predict(fit_PTN, .cols = c(Turbidity_NTU)), PTN_mgperL),
         Silicate_mg.L = ifelse(is.na(Silicate_mg.L), predict(fit_Silicate, .cols = c(Temperature_TBDHereon, Turbidity_NTU, O2_uM, DIC_mg.L, Nitrate_mg.L)), Silicate_mg.L),
         SRP_mgperL = ifelse(is.na(SRP_mgperL), predict(fit_SRP, .cols = c(SRP_mgperL ~ Sat_O2_Perc + TN_mg.L + Chlorophyll_Fluoresence)), SRP_mgperL)) %>% 
  ungroup()

METAG_LF_param_keep <- apply(metadata_METAG_LF, 2, not_any_na)
metadata_METAG_LF <- metadata_METAG_LF[,METAG_LF_param_keep]
#METAT LF
METAT_LF_param_keep <- apply(metadata_METAT_LF, 2, not_any_na)
metadata_METAT_LF <-metadata_METAT_LF[,METAT_LF_param_keep]
#METAG HF
# fit_SRP <- lm(SRP_mgperL ~ Sat_O2_Perc + TN_mg.L + Chlorophyll_Fluoresence, metadata_METAG_HF)
fit_POC <- lm(POC_mgperL  ~  Turbidity_NTU, metadata_METAG_HF)
fit_PTC <- lm(PTC_mgperL  ~  Turbidity_NTU, metadata_METAG_HF)
fit_PTN <- lm(PTN_mgperL  ~  Turbidity_NTU, metadata_METAG_HF)
fit_Silicate <- lm(Silicate_mg.L ~  Temperature_TBDHereon  + Turbidity_NTU + O2_uM + DIC_mg.L +  Nitrate_mg.L, metadata_METAG_HF)

metadata_METAG_HF <- metadata_METAG_HF %>% 
  rowwise() %>% 
  mutate(POC_mgperL = ifelse(is.na(POC_mgperL), predict(fit_POC, .cols = c(Turbidity_NTU)), POC_mgperL),
         PTC_mgperL = ifelse(is.na(PTC_mgperL), predict(fit_PTC, .cols = c(Turbidity_NTU)), PTC_mgperL),
         PTN_mgperL = ifelse(is.na(PTN_mgperL), predict(fit_PTN, .cols = c(Turbidity_NTU)), PTN_mgperL),
         Silicate_mg.L = ifelse(is.na(Silicate_mg.L), predict(fit_Silicate, .cols = c(Temperature_TBDHereon, Turbidity_NTU, O2_uM, DIC_mg.L, Nitrate_mg.L)), Silicate_mg.L),
         SRP_mgperL = ifelse(is.na(SRP_mgperL), predict(fit_SRP, .cols = c(SRP_mgperL ~ Sat_O2_Perc + TN_mg.L + Chlorophyll_Fluoresence)), SRP_mgperL)) %>% 
  ungroup()

METAG_HF_param_keep <- apply(metadata_METAG_HF, 2, not_any_na)
metadata_METAG_HF <- metadata_METAG_HF[,METAG_HF_param_keep]
# METAT HF
fit_POC <- lm(POC_mgperL  ~  Turbidity_NTU, metadata_METAT_HF)
fit_PTC <- lm(PTC_mgperL  ~  Turbidity_NTU, metadata_METAT_HF)
fit_PTN <- lm(PTN_mgperL  ~  Turbidity_NTU, metadata_METAT_HF)

metadata_METAT_HF <- metadata_METAT_HF %>% 
  rowwise() %>% 
  mutate(POC_mgperL = ifelse(is.na(POC_mgperL), predict(fit_POC, .cols = c(Turbidity_NTU)), POC_mgperL),
         PTC_mgperL = ifelse(is.na(PTC_mgperL), predict(fit_PTC, .cols = c(Turbidity_NTU)), PTC_mgperL),
         PTN_mgperL = ifelse(is.na(PTN_mgperL), predict(fit_PTN, .cols = c(Turbidity_NTU)), PTN_mgperL)) %>% 
  ungroup()

METAT_HF_param_keep <- apply(metadata_METAT_HF, 2, not_any_na)
metadata_METAT_HF <- metadata_METAT_HF[,METAT_HF_param_keep]
```

##4. Free-living CCA 
###4.1 Diversity FL 
```{r CCA diversity FL}
#reformat viruses
#remove clusters which sum equal zero where 
#TODO remove cluster originating from samples without metadata Nov 21
sum_FL <- apply(viral_diversity_FL, 1, sum)
t_viral_diversity_FL <- t(viral_diversity_FL[which(sum_FL!=0),])
t_viral_diversity_FL <- as.data.frame(t_viral_diversity_FL)

rownames(t_viral_diversity_FL) <- gsub("GROS22.(1|2)_", "", rownames(t_viral_diversity_FL))
rownames(t_viral_diversity_FL) <- gsub("_METAG.genecount.profile", "", rownames(t_viral_diversity_FL))

cca_diversity_FL <- cca(as.matrix(t_viral_diversity_FL), as.matrix(metadata_METAG_FL))
#select names to display with p.value 0.001
cca_diversity_FL_env <- envfit(cca_diversity_FL, as.matrix(metadata_METAG_FL), permutations = 999)
env_names <-names(cca_diversity_FL_env$vectors$pvals)[which(cca_diversity_FL_env$vectors$pvals==0.001)]
```




```{r}
#make a function for plotting ordination 
ord_plot <- function(ord, color_data, group, arrow_factor=3, manual_colours=c(), env_factors=row.names(as.data.frame(ord$CCA$biplot)), ellipse=F){
veg_1 = as.data.frame(ord$CCA$biplot)
veg_1["env"] = row.names(veg_1)
#importance = variance explained 
ord_sum <- summary(ord)
CCA1_imp <-ord_sum$cont$importance[2,1]*100
CCA2_imp <-ord_sum$cont$importance[2,2]*100
#extracting the data; genus
veg_2 = as.data.frame(ord$CCA$v)
veg_2["primary_cluster"] = row.names(veg_2)
#color data have to contain species indicators (here primary cluster) and grouping data
#group variable points which group will be used for coloring 
veg_2 <- merge(veg_2,color_data, by.x = "row.names", by.y = "primary_cluster", all.x=T)
veg_2 = na.omit(veg_2)

plot = ggplot() +
  geom_point(data =  veg_2, alpha=1, aes(x = CCA1, y = CCA2, colour = veg_2[[group]]))

if (ellipse==T){
plot <- plot + geom_mark_ellipse(data = veg_2, aes(x = CCA1, y = CCA2, colour = veg_2[[group]]))}

veg_1 = veg_1[which(rownames(veg_1) %in% env_factors),]
labels=veg_1$env
labels <- gsub("dCO2_uM", expression("CO"[2]), labels)
labels <- gsub("Temperature_TBDHereon","Temp",labels)
labels <- gsub("TN_mg.L","N",labels)
labels <- gsub("Sat_O2_Perc", expression("%O"[2]), labels)
labels <- gsub("O2_uM", expression("O"[2]), labels)
labels <- gsub("Nitrate_mg.L",expression("NO"[3]), labels)
labels <- gsub("Salinity_PSU","Salinity", labels)
labels <- gsub("Turbidity_NTU ","Turbidity", labels)
labels <- gsub("DOC_mg.L", "DOC", labels)
labels <- gsub("Silicate_mg.L", "Si", labels)

labels <- gsub("TotalDissolvedPhosphate_mg.L", expression("PO"[4]^"3-"), labels)
print(labels)
FL_vir_diversity_plot_class <- plot +
  theme_bw() +
  geom_segment(
    data = veg_1,
    aes(
      x = 0,
      y = 0,
      xend = CCA1 * arrow_factor,
      yend = CCA2 * arrow_factor
    ),
    arrow = arrow(length = unit(0.25, "cm"))
  )   +
  geom_text_repel( parse= TRUE,
    data = veg_1,
    aes(x = CCA1* arrow_factor, y = CCA2 * arrow_factor, label = labels),
    nudge_y=0.2,
    nudge_x=0.2,
    color = "red",
    size = 4
  ) +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18)) + 
  xlab(paste0("CCA1 ", paste0(round(CCA1_imp, 2), "%"))) + 
  ylab(paste0("CCA2 ", paste0(round(CCA2_imp, 2), "%"))) 

if (length(manual_colours)>0){
    FL_vir_diversity_plot_class <- FL_vir_diversity_plot_class + scale_color_manual(values=manual_colours)}
FL_vir_diversity_plot_class
}

```
```{r}
FL_vir_diversity_plot_class <- ord_plot(cca_diversity_FL, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="level_5", env_factors=env_names, ellipse=T)      
FL_vir_diversity_plot_class

tiff("FL_vir_diversity_plot_class_biplot.tiff", unit="px", width = 1000, height = 1000)
FL_vir_diversity_plot_class
dev.off()
```
```{r}

colours=c(brewer.pal(n = length(unique(drep_clusters_fin_taxonomy_unified$Salinity_niche))-2, name = "Blues"), "yellow", "orange")
   
FL_vir_diversity_plot_salinity <- ord_plot(cca_diversity_FL, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Salinity_niche", env_factors=env_names, manual_colours = colours)    
tiff("FL_vir_diversity_salinity_biplot_new.tiff", unit="px", width = 1000, height = 1000)
FL_vir_diversity_plot_salinity + guides(colour = guide_legend(title ="Niche")) 
dev.off()
```

```{r}
FL_vir_diversity_plot_fraction <- ord_plot(cca_diversity_FL, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Fraction_niche", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM"))      
      
FL_vir_diversity_plot_fraction
tiff("FL_vir_diversity_fraction_biplot_new.tiff", unit="px", width = 1000, height = 1000)
FL_vir_diversity_plot_fraction
dev.off()
```


###4.2 Activity FL

```{r CCA activity FL}
sum_FL <- apply(viral_activity_FL, 1, sum)
#reformat viruses
t_viral_activity_FL <- t(viral_activity_FL[which(sum_FL!=0),])
t_viral_activity_FL <- as.data.frame(t_viral_activity_FL)

rownames(t_viral_activity_FL) <- gsub("GROS22.(1|2)_", "", rownames(t_viral_activity_FL))
rownames(t_viral_activity_FL) <- gsub("_METAT.genecount.profile", "", rownames(t_viral_activity_FL))

cca_activity_FL <- cca(as.matrix(t_viral_activity_FL), as.matrix(metadata_METAT_FL))
summary(cca_activity_FL) #scalling of the parameters doesnt matter sqrt of activity results makes proportion explained lower
#select names to display with p.value 0.001
cca_activity_FL_env <- envfit(cca_activity_FL, as.matrix(metadata_METAT_FL), permutations = 999)
env_names <-names(cca_activity_FL_env$vectors$pvals)[which(cca_activity_FL_env$vectors$pvals==0.001)]
```


```{r}
#class
FL_vir_activity_plot_class <- ord_plot(cca_activity_FL, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="level_5", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM"))
FL_vir_activity_plot_class
tiff("FL_vir_activity_plot_class_biplot.tiff", unit="px", width = 1000, height = 1000)
FL_vir_activity_plot_class
dev.off()

#salinity
FL_vir_activity_plot_salinity <- ord_plot(cca_activity_FL, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Salinity_niche", env_factors=env_names, manual_colours = colours)
FL_vir_activity_plot_salinity
tiff("FL_vir_activity_plot_salinity_biplot.tiff", unit="px", width = 1000, height = 1000)
FL_vir_activity_plot_salinity + guides(colour = guide_legend(title ="Niche")) 
dev.off()

#fraction
FL_vir_activity_plot_fraction <- ord_plot(cca_activity_FL, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Fraction_niche", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM"))
FL_vir_activity_plot_fraction
tiff("FL_vir_activity_plot_fraction_biplot.tiff", unit="px", width = 1000, height = 1000)
FL_vir_activity_plot_fraction
dev.off()
```
## 5. Light fraction CCA
### 5.1 Diversity LF  
```{r CCA diversity LF}
#remove rows which sum == 0
sum_LF <- apply(viral_diversity_LF, 1, sum)
#reformat viruses
t_viral_diversity_LF <- t(viral_diversity_LF[which(sum_LF!=0),])
t_viral_diversity_LF <- as.data.frame(t_viral_diversity_LF)
ps <- 0.000001

t_viral_diversity_LF <- t_viral_diversity_LF 
rownames(t_viral_diversity_LF) <- gsub("GROS22.(1|2)_", "", rownames(t_viral_diversity_LF))
rownames(t_viral_diversity_LF) <- gsub("_METAG.genecount.profile", "", rownames(t_viral_diversity_LF))

cca_diversity_LF <- cca(as.matrix(t_viral_diversity_LF), as.matrix(metadata_METAG_LF))
summary(cca_diversity_LF) #scalling of the parameters doesnt matter sqrt of diversity results makes proportion explained lower
#select names to display with p.value 0.001
cca_diversity_LF_env <- envfit(cca_diversity_LF, as.matrix(metadata_METAG_LF), permutations = 999)
env_names <-names(cca_diversity_LF_env$vectors$pvals)[which(cca_diversity_LF_env$vectors$pvals==0.001)]
```


```{r}
#class
LF_vir_diversity_plot_class <- ord_plot(cca_diversity_LF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="level_5", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM"))
LF_vir_diversity_plot_class
tiff("LF_vir_diversity_plot_class_biplot.tiff", unit="px", width = 1000, height = 1000)
LF_vir_diversity_plot_class
dev.off()

#fraction
LF_vir_diversity_plot_fraction_niche <- ord_plot(cca_diversity_LF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Fraction_niche", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM"))
LF_vir_diversity_plot_fraction_niche

tiff("LF_vir_diversity_plot_fraction_biplot.tiff", unit="px", width = 1000, height = 1000)
LF_vir_diversity_plot_fraction_niche
dev.off()

#salinity
LF_vir_diversity_plot_salinity <- ord_plot(cca_diversity_LF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Salinity_niche", env_factors=env_names, manual_colours = colours) 
LF_vir_diversity_plot_salinity  
tiff("LF_vir_diversity_plot_salinity_biplot.tiff", unit="px", width = 1000, height = 1000)
LF_vir_diversity_plot_salinity + guides(colour = guide_legend(title ="Niche")) 
dev.off()

```
### 5.2 Activity LF

```{r CCA activity FL}
#remove rows which sum == 0
sum_LF <- apply(viral_activity_LF, 1, sum)
#reformat viruses
t_viral_activity_LF <- t(viral_activity_LF[which(sum_LF!=0),])
t_viral_activity_LF <- as.data.frame(t_viral_activity_LF)

rownames(t_viral_activity_LF) <- gsub("GROS22.(1|2)_", "", rownames(t_viral_activity_LF))
rownames(t_viral_activity_LF) <- gsub("_METAT.genecount.profile", "", rownames(t_viral_activity_LF))

cca_activity_LF <- cca(as.matrix(t_viral_activity_LF), as.matrix(metadata_METAT_LF))
summary(cca_activity_LF) #scalling of the parameters doesnt matter sqrt of activity results makes proportion explained lower
#select names to display with p.value 0.001
cca_activity_LF_env <- envfit(cca_activity_LF, as.matrix(metadata_METAT_LF), permutations = 999)
env_names <-names(cca_activity_LF_env$vectors$pvals)[which(cca_activity_LF_env$vectors$pvals==0.001)]
```

```{r}
#class
LF_vir_activity_plot_class <-  ord_plot(cca_activity_LF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="level_5", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM"))  
LF_vir_activity_plot_class
tiff("LF_vir_activity_plot_class_biplot.tiff", unit="px", width = 1000, height = 1000)
LF_vir_activity_plot_class
dev.off()

#fraction
LF_vir_activity_plot_fraction_niche <- ord_plot(cca_activity_LF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Fraction_niche", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM"))  
LF_vir_activity_plot_fraction_niche

tiff("LF_vir_activity_plot_fraction_biplot.tiff", unit="px", width = 1000, height = 1000)
LF_vir_activity_plot_fraction_niche
dev.off()

#salinity
LF_vir_activity_plot_salinity <-  ord_plot(cca_activity_LF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Salinity_niche", env_factors=env_names, manual_colours = colours)
LF_vir_activity_plot_salinity  
tiff("LF_vir_activity_plot_salinity_biplot.tiff", unit="px", width = 1000, height = 1000)
LF_vir_activity_plot_salinity
dev.off()
```


##6. Heavy fraction CCA
### 6.1. Diversity HF
```{r CCA diversity HF}
sum_LF <- apply(viral_diversity_HF, 1, sum)
#reformat viruses
t_viral_diversity_HF <- t(viral_diversity_HF[which(sum_LF!=0),])
t_viral_diversity_HF <- as.data.frame(t_viral_diversity_HF)

rownames(t_viral_diversity_HF) <- gsub("GROS22.(1|2)_", "", rownames(t_viral_diversity_HF))
rownames(t_viral_diversity_HF) <- gsub("_METAG.genecount.profile", "", rownames(t_viral_diversity_HF))

cca_diversity_HF <- cca(as.matrix(t_viral_diversity_HF), as.matrix(metadata_METAG_HF))
summary(cca_diversity_HF) #scalling of the parameters doesnt matter sqrt of diversity results makes proportion explained lower
#select names to display with p.value 0.001
cca_diversity_HF_env <- envfit(cca_diversity_HF, as.matrix(metadata_METAG_HF), permutations = 999)
env_names <-names(cca_diversity_HF_env$vectors$pvals)[which(cca_diversity_HF_env$vectors$pvals==0.001)]

```


```{r}
#class
HF_vir_diversity_plot_class <- ord_plot(cca_diversity_HF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="level_5", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM"))  
HF_vir_diversity_plot_class
tiff("HF_vir_diversity_plot_class_biplot.tiff", unit="px", width = 1000, height = 1000)
HF_vir_diversity_plot_class
dev.off()

#fraction
HF_vir_diversity_plot_fraction_niche <-ord_plot(cca_diversity_HF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Fraction_niche", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM"))  
HF_vir_diversity_plot_fraction_niche

tiff("HF_vir_diversity_plot_fraction_biplot.tiff", unit="px", width = 1000, height = 1000)
HF_vir_diversity_plot_fraction_niche
dev.off()

#salinity
HF_vir_diversity_plot_salinity <- ord_plot(cca_diversity_HF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Salinity_niche", env_factors=env_names, manual_colours = colours) 
HF_vir_diversity_plot_salinity  
tiff("HF_vir_diversity_plot_salinity_biplot.tiff", unit="px", width = 1000, height = 1000)
HF_vir_diversity_plot_salinity
dev.off()

```

Salinity and temperature looks prominent.


###6.2 Activity HF

```{r CCA activity HF}
sum_LF <- apply(viral_activity_HF, 1, sum)
#reformat viruses
t_viral_activity_HF <- t(viral_activity_HF[which(sum_LF!=0),])
t_viral_activity_HF <- as.data.frame(t_viral_activity_HF)

rownames(t_viral_activity_HF) <- gsub("GROS22.(1|2)_", "", rownames(t_viral_activity_HF))
rownames(t_viral_activity_HF) <- gsub("_METAT.genecount.profile", "", rownames(t_viral_activity_HF))

cca_activity_HF <- cca(as.matrix(t_viral_activity_HF), as.matrix(metadata_METAT_HF))
summary(cca_activity_HF) #scalling of the parameters doesnt matter sqrt of activity results makes proportion explained lower

#select names to display with p.value 0.001
cca_activity_HF_env <- envfit(cca_activity_HF, as.matrix(metadata_METAT_HF), permutations = 999)
env_names <-names(cca_activity_HF_env$vectors$pvals)[which(cca_activity_HF_env$vectors$pvals==0.001)]

```



```{r}
#class
HF_vir_activity_plot_class <- ord_plot(cca_activity_HF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="level_5", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM")) 
HF_vir_activity_plot_class
tiff("HF_vir_activity_plot_class_biplot.tiff", unit="px", width = 1000, height = 1000)
HF_vir_activity_plot_class
dev.off()

#fraction
HF_vir_activity_plot_fraction_niche <- ord_plot(cca_activity_HF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Fraction_niche", env_factors=c("Phosphate_uM", "Nitrate_uM", "pH_TBDHereon", "Salinity_PSU", "Temperature_TBDHereon", "Turbidity_NTU", "Total_DIN_uM")) 
HF_vir_activity_plot_fraction_niche

tiff("HF_vir_activity_plot_fraction_biplot.tiff", unit="px", width = 1000, height = 1000)
HF_vir_activity_plot_fraction_niche
dev.off()

#salinity
HF_vir_activity_plot_salinity <- ord_plot(cca_activity_HF, color_data=drep_clusters_fin_taxonomy_unified[, c("primary_cluster", "level_5", "Salinity_niche", "Fraction_niche")],  group="Salinity_niche", env_factors=env_names, manual_colours = colours)
HF_vir_activity_plot_salinity  
tiff("HF_vir_activity_plot_salinity_biplot.tiff", unit="px", width = 1000, height = 1000)
HF_vir_activity_plot_salinity
dev.off()
```


```{r plot cca all}
library(ggpubr)
A <-FL_vir_diversity_plot_salinity  + guides(colour="none")
B <-LF_vir_diversity_plot_salinity + guides(colour="none")
C <-HF_vir_diversity_plot_salinity + theme(legend.text=element_text(size=12), legend.title=element_text(size=15)) + guides(colour = guide_legend(title ="Niche")) 
D <-FL_vir_activity_plot_salinity + guides(colour="none")
E <-LF_vir_activity_plot_salinity + guides(colour="none")
Fa <-HF_vir_activity_plot_salinity + guides(colour="none")
grid <- ggarrange(A, B, C, D, E, Fa, ncol = 3, nrow=2, legend="bottom", common.legend = TRUE) 

tiff("all_CCA_3.tiff", unit="px", width = 1000, height = 1000)
grid
dev.off()
```


##8. Build models
Building models for salinity and phosphate predictions

We will try to use viral abundance data based on individual marker genes 
```{r salinity }
library(glmnet)
library(caret)
#divide dataset by test and train
#try METAG first
#we have to transform data so we have counts of each primary cluster in separate column and each row is a different samples with a parameter
geneabund_drep_marker_taxonomy_long <- readRDS(file="geneabund_drep_marker_taxonomy_long.RDS")
#remove taxonomy distribution so we dont we multiple separe vectors for the same sample but one which represents community
#keep only one replicate - keep only even Associated numbers
geneabund_drep_marker_taxonomy_wide <-  geneabund_drep_marker_taxonomy_long %>% ungroup() %>% filter(Associatednumber%%2==0) %>% select(-c("lineage", "level_5", "Salinity_niche", "Fraction_niche"))  %>%  pivot_wider(names_from=primary_cluster, values_from=counts, values_fill = 0, names_prefix="cluster_") %>% distinct()
geneabund_drep_marker_taxonomy_wide_METAG <- geneabund_drep_marker_taxonomy_wide %>% filter(data_type == "METAG") 
                                                                                            
#keep rows where we have collected param
geneabund_drep_marker_taxonomy_wide_METAG <- geneabund_drep_marker_taxonomy_wide_METAG %>% filter(!is.na(Salinity_PSU))

geneabund_drep_marker_taxonomy_wide_METAT <- geneabund_drep_marker_taxonomy_wide %>% filter(data_type == "METAT")
#draw randomly train and test set 80:20
set.seed(123)
p <- nrow(geneabund_drep_marker_taxonomy_wide_METAG)
train_rows <- sample(size=p*0.8, 1:p)
geneabund_drep_marker_taxonomy_wide_METAG_train <- geneabund_drep_marker_taxonomy_wide_METAG[train_rows,]
geneabund_drep_marker_taxonomy_wide_METAG_train_x <- geneabund_drep_marker_taxonomy_wide_METAG_train %>% ungroup() %>% select(starts_with("cluster"))
geneabund_drep_marker_taxonomy_wide_METAG_test <- geneabund_drep_marker_taxonomy_wide_METAG[-train_rows,]
geneabund_drep_marker_taxonomy_wide_METAG_test_x <- geneabund_drep_marker_taxonomy_wide_METAG_test %>% ungroup() %>% select(starts_with("cluster"))
y_train <- geneabund_drep_marker_taxonomy_wide_METAG_train$Salinity_PSU 
y_test <- geneabund_drep_marker_taxonomy_wide_METAG_test$Salinity_PSU 
# #try the same for phosphate 
# y_train <- as.vector(geneabund_drep_marker_taxonomy_wide_METAG_train$Phosphate_uM) 
# y_test <- as.vector(geneabund_drep_marker_taxonomy_wide_METAG_test$Phosphate_uM)

```
We will try to use viral abundance data based on individual marker genes 
```{r model_salinity}
library(caret)
#divide dataset by test and train
#try METAG first
#we have to transform data so we have counts of each primary cluster in separate column and each row is a different samples with a parameter
geneabund_drep_marker_taxonomy_long <- readRDS(file="geneabund_drep_marker_taxonomy_long.RDS")
#remove taxonomy distribution so we dont we multiple separe vectors for the same sample but one which represents community
geneabund_drep_marker_taxonomy_wide <-  geneabund_drep_marker_taxonomy_long %>% ungroup() %>% select(-c("lineage", "level_5"))  %>% filter(Associatednumber%%2==0) %>%  pivot_wider(names_from=primary_cluster, values_from=counts, values_fill = 0, names_prefix="cluster_") %>% distinct()
#remove outliers - our model can be used for fresh to mesohaline water
geneabund_drep_marker_taxonomy_wide_METAG_salinity <- geneabund_drep_marker_taxonomy_wide %>% filter(data_type == "METAG") %>% filter(Salinity_PSU < 17.0) 
set.seed(123)
#divide by train and test set to get balanced datasets
trainIndex <- createDataPartition(geneabund_drep_marker_taxonomy_wide_METAG$Salinity_level, p = .8, 
                                  list = FALSE, 
                                  times = 1)
virusTrain <- geneabund_drep_marker_taxonomy_wide_METAG[ trainIndex,]
virusTest  <- geneabund_drep_marker_taxonomy_wide_METAG[-trainIndex,]

set.seed(123)
#divide train set to folds
train_control <- trainControl(method = "cv",
                              number=5,
                              savePredictions = TRUE,
                              verboseIter = TRUE,
                              allowParallel = FALSE)
#generate formula so I dont need to generate new data frame for each model - saving memory

#find columns with near zero variance 0 take into account problematic fold 
colzerovar<- nearZeroVar(virusTrain)
# #and remove the m from formula
# 
clusters_model <- colnames(virusTrain)[-c(colzerovar)]
#take column names with clusters with more then near zero variance for final model
clusters_model <- clusters_model[grepl("cluster",clusters_model)]
#take part of the data as it gives error when trying to create long formula
#search space

grid <- expand.grid(
  alpha = seq(0, 1, length = 100),
  lambda = seq(0, 10, length = 100)
)
model_salinity <- train(x=as.matrix(virusTrain[clusters_model]), y=as.numeric(virusTrain$Salinity_PSU), 
               method = "glmnet", na.action = na.omit,  #preProc = c("center", "scale"),
               trControl = train_control, tuneGrid = grid, verbose = TRUE,  allowParallel=F,
               importance=T, metric="Rsquared")
# save.image(file="cca_individual_marker_gene.RData")

print(model_salinity)

best_model_stats <- model_salinity$results[which(model_salinity$results$alpha == model_salinity$bestTune$alpha & model_salinity$results$lambda == model_salinity$bestTune$lambda ),]
cv_eval <- model_salinity$resample
```

```{r eval_salinity}
pred<-predict(object=model_salinity, newdata = as.matrix(virusTest[,clusters_model]))

plot(virusTest$Salinity_PSU, pred, ylim=c(0,20))
lm_pred<-lm(virusTest$Salinity_PSU ~ pred, b=1)
abline(lm_pred, col=c("red"))
abline(a=0, b=1, col ="black")
summary(lm_pred)


r2 <- summary(lm_pred)$r.squared
tiff("model_salinity_validation2.tiff")
pred_validation <- ggplot() + geom_point(aes(x=virusTest$Salinity_PSU, y=pred)) + geom_abline(color="red", size=1) + geom_smooth(aes(x=virusTest$Salinity_PSU, y=pred), method = "lm", se = FALSE) +  annotate(geom="text", x=7.5, y=15, label=paste("Rsquared = ",round(r2,2)), size= 10, colour = "#3366FF") + ylab("Predicted PSU") + xlab("Observed PSU") + theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20)) + theme_bw(base_size = 20)
pred_validation
dev.off()
#looks like two groups - 
#try to find wich samples have the biggest mae ~ 7 - check if they have sth in common

```

```{r model_phosphate}
library(caret)
#divide dataset by test and train
#try METAG first
#we have to transform data so we have counts of each primary cluster in separate column and each row is a different samples with a parameter
# geneabund_drep_marker_taxonomy_long <- readRDS(file="geneabund_drep_marker_taxonomy_long.RDS")
# #remove taxonomy distribution so we dont we multiple separe vectors for the same sample but one which represents community
# geneabund_drep_marker_taxonomy_wide <-  geneabund_drep_marker_taxonomy_long %>% ungroup() %>% select(-c("lineage", "level_5"))  %>%  pivot_wider(names_from=primary_cluster, values_from=counts, values_fill = 0, names_prefix="cluster_") %>% distinct()
# geneabund_drep_marker_taxonomy_wide_METAG <- geneabund_drep_marker_taxonomy_wide %>% filter(data_type == "METAG") 
#reshuffle data and draw folds
set.seed(123)
#divide by train and test set to get balanced datasets
geneabund_drep_marker_taxonomy_wide_METAG_phosphate <- geneabund_drep_marker_taxonomy_wide_METAG[which(!is.na(geneabund_drep_marker_taxonomy_wide_METAG$Phosphate_uM)), ]
trainIndex <- createDataPartition(geneabund_drep_marker_taxonomy_wide_METAG_phosphate$Phosphate_uM, p = .8,
                                  list = FALSE,
                                  times = 1)
virusTrain <- geneabund_drep_marker_taxonomy_wide_METAG_phosphate[ trainIndex,]
virusTest  <- geneabund_drep_marker_taxonomy_wide_METAG_phosphate[-trainIndex,]

set.seed(123)
#divide train set to folds
# folds <- createFolds(y=as.factor(virusTrain$Salinity_level), k = 5)
train_control <- trainControl(method = "cv",
                              number=5,
                              savePredictions = TRUE,
                              verboseIter = TRUE,
                              allowParallel = FALSE)
#generate formula so I dont need to generate new data frame for each model - saving memory

#find columns with near zero variance 0 take into account problematic fold 
colzerovar<- nearZeroVar(virusTrain)
# #and remove the m from formula
# 
clusters_model <- colnames(virusTrain)[-c(colzerovar)]
#take column names with clusters with more then near zero variance for final model
clusters_model <- clusters_model[grepl("cluster",clusters_model)]
#take part of the data as it gives error when trying to create long formula
model_Phosphate <- train(x=as.matrix(virusTrain[clusters_model]), y=as.numeric(virusTrain$Phosphate_uM), 
               method = "glmnet", na.action = na.omit,  #preProc = c("center", "scale"),
               trControl = train_control, tuneLength = 10, verbose = TRUE,  allowParallel=F,
               importance=T, metric="Rsquared")
# save.image(file="cca_individual_marker_gene")

print(model_Phosphate)

best_model_stats <- model_Phosphate$results[which(model_Phosphate$results$alpha == model_Phosphate$bestTune$alpha & model_Phosphate$results$lambda == model_Phosphate$bestTune$lambda ),]
cv_eval <- model_Phosphate$resample
```
```{r eval_phosphate}
pred<-predict(object=model_Phosphate, newdata = as.matrix(virusTest[,clusters_model]))

plot(virusTest$Phosphate_uM, pred, ylim=c(0,5))
lm_pred<-lm(virusTest$Phosphate_uM ~ pred, b=1)
abline(lm_pred, col=c("red"))
abline(a=0, b=1, col ="black")
summary(lm_pred)


r2 <- summary(lm_pred)$r.squared
model_Phosphate_validation.tiff<-ggplot() + geom_point(aes(x=virusTest$Phosphate_uM, y=pred)) + geom_abline(color="red", size=1) + geom_smooth(aes(x=virusTest$Phosphate_uM, y=pred), method = "lm", se = FALSE) +  annotate(geom="text", x=1.5, y=15, label=paste("Rsquared = ",round(r2,2)), size= 10, colour = "#3366FF") + ylab("Predicted") + xlab("Observed") + theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))
tiff("model_Phosphate_validation.tiff")
model_Phosphate_validation.tiff
dev.off()
#looks like two groups - 
#try to find which samples have the biggest mae ~ 7 - check if they have sth in common

```
```{r model_temp}
library(caret)
#divide dataset by test and train
#try METAG first

#reshuffle data and draw folds
set.seed(123)
#divide by train and test set to get balanced datasets
geneabund_drep_marker_taxonomy_wide_METAG <- geneabund_drep_marker_taxonomy_wide %>% filter(data_type == "METAG")

trainIndex <- createDataPartition(geneabund_drep_marker_taxonomy_wide_METAG$Temperature_TBDHereon, p = .8, list = FALSE,  times = 1)
virusTrain <- geneabund_drep_marker_taxonomy_wide_METAG[ trainIndex,]
virusTest  <- geneabund_drep_marker_taxonomy_wide_METAG[-trainIndex,]

set.seed(123)
#divide train set to folds
# folds <- createFolds(y=as.factor(virusTrain$Salinity_level), k = 5)
train_control <- trainControl(method = "cv",
                              number=5,
                              savePredictions = TRUE,
                              verboseIter = TRUE,
                              allowParallel = FALSE)
#generate formula so I dont need to generate new data frame for each model - saving memory

#find columns with near zero variance 0 take into account problematic fold 
colzerovar<- nearZeroVar(virusTrain)
# #and remove the m from formula
# 
clusters_model <- colnames(virusTrain)[-c(colzerovar)]
#take column names with clusters with more then near zero variance for final model
clusters_model <- clusters_model[grepl("cluster",clusters_model)]
#take part of the data as it gives error when trying to create long formula
model_temp <- train(x=as.matrix(virusTrain[clusters_model]), y=as.numeric(virusTrain$Temperature_TBDHereon), 
               method = "glmnet", na.action = na.omit,  #preProc = c("center", "scale"),
               trControl = train_control, tuneLength = 10, verbose = TRUE,  allowParallel=F,
               importance=T, metric="Rsquared")
# save.image(file="cca_individual_marker_gene")

print(model_temp)

best_model_stats <- model_temp$results[which(model_temp$results$alpha == model_temp$bestTune$alpha & model_temp$results$lambda == model_temp$bestTune$lambda ),]
cv_eval <- model_temp$resample
```
```{r eval_temp}
pred<-predict(object=model_temp, newdata = as.matrix(virusTest[,clusters_model]))

plot(virusTest$Temperature_TBDHereon, pred)
lm_pred<-lm(virusTest$Temperature_TBDHereon ~ pred, b=1)
abline(lm_pred, col=c("red"))
abline(a=0, b=1, col ="black")
summary(lm_pred)


r2 <- summary(lm_pred)$r.squared
model_temp_validation<-ggplot() + geom_point(aes(x=virusTest$Temperature_TBDHereon, y=pred)) + geom_abline(color="red", size=1) + geom_smooth(aes(x=virusTest$Temperature_TBDHereon, y=pred), method = "lm", se = FALSE) +  annotate(geom="text", x=8, y=20, label=paste("Rsquared = ",round(r2,2)), size= 10, colour = "#3366FF") + ylab("Predicted") + xlab("Observed" ) + ylim(c(0,25)) + theme(axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20)) + theme_bw(base_size = 20)
tiff("model_temp_validation.tiff")
model_temp_validation
dev.off()
#looks like two groups - 
#try to find which samples have the biggest mae ~ 7 - check if they have sth in common

```
## 9. Viral diversity indexes - shannon simpson, richness based on an occurence and niches
```{r}
geneabund_drep_marker_taxonomy_long <- geneabund_drep_marker_taxonomy_long %>% merge(drep_clusters_fin_taxonomy_unified[,c("primary_cluster", "Salinity_niche","Fraction_niche") ], by="primary_cluster")
# permanova_salinity_niche <- adonis2(geneabund_drep_marker_taxonomy_wide[, grepl("cluster", colnames(geneabund_drep_marker_taxonomy_wide))] ~ Salinity_niche, data =geneabund_drep_marker_taxonomy_long, permutations = 999)
# permanova_salinity_level <- adonis2(formula = permanova_salinity_level_pairwise[, grepl("cluster", colnames(permanova_salinity_level_pairwise))] ~ Salinity_level, data = permanova_salinity_level_pairwise, permutations = 999)

library(pairwiseAdonis)
permanova_salinity_level_pairwise <- pairwise.adonis(geneabund_drep_marker_taxonomy_wide_METAG[, grepl("cluster", colnames(geneabund_drep_marker_taxonomy_wide_METAG))], geneabund_drep_marker_taxonomy_wide_METAG$Salinity_level)
# permanova_fraction_niche <- adonis2(formula = geneabund_drep_marker_taxonomy_wide[$counts] ~ Fraction_niche, data = geneabund_drep_marker_taxonomy_long, permutations = 999)
```
```{r diverity meassurments}
library(phyloseq)
library(ggpubr)
otu1 <- otu_table(geneabund_drep_marker_taxonomy_wide_METAG[, grepl("cluster", colnames(geneabund_drep_marker_taxonomy_wide_METAG))], taxa_are_rows = F)
geneabund_drep_marker_taxonomy_wide_METAG$Salinity_level <- factor(geneabund_drep_marker_taxonomy_wide_METAG$Salinity_level, levels=c("freshwater", "oligohaline", "mesohaline", "polyhaline" ))
phyloseq <- phyloseq(otu1, my_sample_data=sample_data(geneabund_drep_marker_taxonomy_wide_METAG))
a_my_comparisons <- list(c("oligohaline", "freshwater"), c("oligohaline", "mesohaline"), c("oligohaline","polyhaline"), c("freshwater", "mesohaline"), c("freshwater", "polyhaline"), c("mesohaline", "polyhaline"))
richness <- estimate_richness(phyloseq, measures=c("Simpson", "Shannon")) 
observed_otu<-100*geneabund_drep_marker_taxonomy_wide_METAG[, grepl("cluster", colnames(geneabund_drep_marker_taxonomy_wide_METAG))]
otu2 <- otu_table(observed_otu, taxa_are_rows = F)

phyloseq2 <- phyloseq(otu2, my_sample_data=sample_data(geneabund_drep_marker_taxonomy_wide_METAG))
richness2 <- estimate_richness(phyloseq2, measures=c("Observed")) 
symnum<- list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))
#all alpha diversity meassurments

tiff("Alpha_diversity_Salinity.tiff",units = 'px', width = 1000, height = 500)
plot_richness(phyloseq2, x="Salinity_level", measures=c( "Observed", "Simpson", "Shannon")) + geom_boxplot(fill=rep(colors,3))  + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum) + xlab("Salinity") + theme_bw(base_size=20) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(labels = c("Freshwater", "Oligohaline",  "Mesohaline", "Polyhaline"))
dev.off()

kruskal.test(richness$Shannon ~ sample_data(phyloseq)$Salinity_level)
kruskal.test(richness$Simpson ~ sample_data(phyloseq)$Salinity_level)
library(FSA)
dunn_shannon_salinity <- dunnTest(Shannon ~ sample_data(phyloseq)$Salinity_level,
                  data=richness,
                  method="bh")
dunn_shannon_fraction <- dunnTest(Shannon ~ sample_data(phyloseq)$Sample_type,
                  data=richness,
                  method="bh")

plot_richness(phyloseq, x="Salinity_level", measures=c( "Simpson")) + geom_boxplot() + geom_pwc(method ="dunn_test")
# +stat_pvalue_manual(dunn, label = "p = {scales::pvalue(p.adj)}")
symnum<- list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))
plot_richness_shannon <- plot_richness(phyloseq, x="Salinity_level", measures=c( "Shannon")) + geom_boxplot(fill=colors) + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum) + xlab("Salinity level") 
tiff("plot_richness_shannon.tiff")
plot_richness_shannon
dev.off()

plot_richness_simpson <- plot_richness(phyloseq, x="Salinity_level", measures=c("Simpson")) + geom_boxplot() + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum) + xlab("Salinity level") 
plot_richness(phyloseq, x="Station", measures=c( "Simpson")) + geom_boxplot() + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum)
plot_richness(phyloseq, x="Sample_type", measures=c( "Simpson")) + geom_boxplot()  + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum) 

plot_richness(phyloseq, x="Sample_date", measures=c( "Simpson")) + geom_boxplot()  + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum)
plot_richness(phyloseq, x="Salinity_level", measures=c( "Shannon")) + geom_boxplot()  + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum) 
plot_richness(phyloseq, x="Station", measures=c( "Shannon")) + geom_boxplot()  + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum)
plot_richness(phyloseq, x="Sample_type", measures=c( "Shannon")) + geom_boxplot()  + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum) 
plot_richness(phyloseq, x="Sample_date", measures=c( "Shannon")) + geom_boxplot() + geom_pwc(method ="dunn_test", hide.ns ="p", label = "p.signif", symnum.args=symnum)
```
