---
title: "Phototrophy analysis"
author: "Julia Golebiowska"
date: "2024-03-14"
output: pdf_document
---
 
```{r include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
setwd("~/IGB_phd/phototrophy")
path = "C:/Users/jgolebiowska/Documents/IGB_phd/workshop_gene_catalog"
```

```{r read , eval=FALSE, include=FALSE}
##Loading files
#gene_cluster_table
'''
clstr <- read.csv(paste0(path,"/GROS22.clstr"), sep="\t", skip=0, header = F)
colnames(clstr) <- c("gene_cluster", "mem", "gene", "length", "pos")
#annotation
annot <- read.csv(paste0(path,"/GROS22.kegg-annotations.tsv"), sep="\t", skip=0, header = TRUE, quote="")
#abundance
geneabund <- read.csv(paste0(path,"/BICEST_cellabund.tsv"), sep="\t", skip=1, header = TRUE, row.names = 1)
#prepare geneabund for merging  
geneabund <- geneabund %>% 
  rownames_to_column("gene_cluster")
#prepare metadata from merging
metadata_sim <- read.csv(paste0(path,"/SAMEAID_SampleID_simplified.csv"), header=TRUE, sep=";") 
#select the needed columns to avoid doubling computations
metadata_sim <-metadata_sim[,c("BioSample", "ProjectID")]
#read simplified metadata, as it has a ProjectID 
metadata <- read.csv(paste0(path,"/SAMEAID_SampleID.csv"), header=TRUE, sep=";") %>% 
  left_join(metadata_sim, by = "BioSample")  %>% 
  mutate(sampleid=paste0(ProjectID,"_",BioSample,"_METAT.genecount.profile")) %>%
  mutate(data_type="METAT")

metadata2 <- read.csv(paste0(path, "/SAMEAID_SampleID.csv"), header=TRUE, sep=";") %>% 
  left_join(metadata_sim, by = "BioSample")  %>% 
  mutate(sampleid=paste0(ProjectID,"_",BioSample,"_METAG.genecount.profile")) %>%
  mutate(data_type="METAG")

metadata <- rbind(metadata, metadata2)
save.image(file="phototrophy.RData")

'''
```

```{r load}
#or just load prepared initial data
load(file="phototrophy.RData")
```
# 1. Rhodopsins analysis

## 1.1. Find rhodopsins genes
```{r, echo=FALSE}
#let's find rhodopsins using KO
#here the ids from KEGG database
#K00909 GRK1_7; rhodopsin kinase [EC:2.7.11.14]
#K04641 bop; bacteriorhodopsin
#K04642 hop; halorhodopsin
#K04643 sop; sensory rhodopsin
rhd_clust <- annot %>%  
  filter(KO %in% c("K00909","K04641","K04642","K04643")) %>% 
  left_join(geneabund,  join_by(QUERY == gene_cluster))
#not enough to do meaningful analysis 
#interestigly all rhodopsins were found in viral scaffolds ?
#only sensory rhodopsins - so rather not realy responsible for phototrophy
unique(rhd_clust[,"KO"])
```
## 1.2. Explore rhodopsins spatial and temporal distribution
```{r}
#prepare the table with matadata for plot
meta_rhd_clust <- rhd_clust %>% 
  pivot_longer(!c("QUERY", "KO", "BRITE", "CAZY", "COG", "DISEASE", "DRUG", "ENZYME", "GO", "MODULE", "NETWORK", "PATHWAY", "PUBMED", "RCLASS", "REACTION", "TC", "VP", "DESCRIPTION", "length"), names_to = "sampleid", values_to = "counts", values_drop_na = 0) %>%
  merge(metadata)
meta_rhd_clust$Stromkilometer <- sub(",", ".", meta_rhd_clust$Stromkilometer)
meta_rhd_clust$Stromkilometer <- as.double(meta_rhd_clust$Stromkilometer)

#plot it!
meta_rhd_clust %>%  
  filter(Station != 'BunthausSpitze') %>% 
  filter(Sample_date != "Nov 21") %>%
  mutate(counts=counts/1000) %>%
  ggplot(aes(x=Stromkilometer, y=counts)) + geom_point() +
  geom_tile() + facet_wrap(~data_type *Sample_date) + ggtitle("Rhd counts across different seasons, river sites and data types") + geom_smooth(method = lm)
#not enough to do meaningful analysis 
#interestigly all rhodopsins were found in viral scaffolds ?
#usually the highest variance in the last station
```
## 1.3. Check different fractions
```{r}
#check different fractions
#plot it!
meta_rhd_clust %>%  
  filter(Station != 'BunthausSpitze') %>% 
  filter(Sample_date != "Nov 21") %>%
  mutate(counts=counts/1000) %>%
  ggplot(aes(x=Stromkilometer, y=counts)) + geom_point() +
  geom_tile()  + facet_wrap(~ Sample_type) 
#higher variance in free living
```
## 1.4. Check correlations
```{r}
#explore correlations for METAT counts
numeric_col <-  c("counts","Stromkilometer", "Sat_O2_TBDHereon","Turbidity_TBDHereon", "SPM_mgperL", "DOC_mg.L","DIC_mg.L","Temperature_TBDHereon", "pH_TBDHereon", "Silicate_mg.L", "Nitrite_mg.L", "TotalDissolvedPhosphate_mg.L", "RespirationRate_O2ug.L.h",  "Phosphate_µM", "O2_TBDHereon", "TN_mg.L", "DIC_uM.L", "DOC_uM.L",  "POC_mgperL", "PTH_mgperL", "Ammonium_mg.L", "SRP_µM", "Salinity_TBDHereon", "PTH_mgperL", "POC_mgperL", "PTC_mgperL", "PTN_mgperL")
meta_rhd_clust_df <-  meta_rhd_clust  %>% filter(data_type == "METAT") %>%
  filter(Station != 'BunthausSpitze') %>% 
  filter(Sample_date != "Nov 21") %>% 
  filter(data_type == 'METAT') %>% 
  select(numeric_col) 
 
meta_rhd_clust_df[ meta_rhd_clust_df == ""] <- NA 

for (i in c(3:length(meta_rhd_clust_df[1,]))){
  meta_rhd_clust_df[,i] <-  as.numeric(sub(",", ".", meta_rhd_clust_df[,i])) #replace , to . to get a string which can be converted to double 
}
meta_rhd_clust_df[is.na(meta_rhd_clust_df)] <- 0 
head(meta_rhd_clust_df)
#to numeric 
meta_rhd_clust_matrix <- matrix(apply(meta_rhd_clust_df,2, as.numeric), ncol=ncol(meta_rhd_clust_df), dimnames=list(c(),colnames(meta_rhd_clust_df)))

meta_rhd_clust_matrix[is.na(meta_rhd_clust_matrix)] <- 0 
res <- cor(meta_rhd_clust_matrix)
res[ is.na(res)] <- 0 
heatmap(round(res, 2), cexRow = 0.5, cexCol = 0.5)
```
Very strong correlation between "DOC_mg.L", "TN_mg.L". Quite strong between  "TN_mg.L", "DIC_mg.L"
## 1.5. Try to make a model
```{r}

#lets plot two parameters agains rhodopsins counts
#salinity
salinitymodel <- lm(counts ~ Salinity_TBDHereon, data = meta_rhd_clust_df) # didn't work

meta_rhd_clust_df %>% 
ggplot(aes(x=Salinity_TBDHereon, y=counts)) + geom_point() + geom_line(aes(x= Salinity_TBDHereon , y = Salinity_TBDHereon*salinitymodel$coefficients[2]+salinitymodel$coefficients[1] )) 

#pH
ggplot(meta_rhd_clust_df[meta_rhd_clust_df["pH_TBDHereon"]>0,], aes(x=pH_TBDHereon, y=log(counts))) + geom_point() + geom_smooth() #ommit rows with zero - artifact from NA <= 0
```
Very subtle tendency
## 1.6. Explore variance
```{r}
#explore the variance as it looks like is different across stations
var_meta_rhd_clust_station <- meta_rhd_clust %>%  
  filter(Station != 'BunthausSpitze') %>% 
  filter(Sample_date != "Nov 21") %>% 
  group_by(Station) %>% 
  summarise(var = var(counts))

#variance across stations (similarly the variance is lower n the brakish water)
rhd_var_station <- var_meta_rhd_clust_station  %>% arrange(factor(Station, levels =c("Muehlenberger Loch","Twielenfleth", "Schwarztonnensand", "Brunsbuettel","Meedem Grund")))

rhd_var_station #change the order
```

# 2. PsbA analysis 
## 2.1 Find PsbA gene and check its spatial and temporal distribution

```{r}
#find psbA genes 
psbA_clust <- annot %>%  
  filter(KO %in% c("K02703")) %>% 
  left_join(geneabund,  join_by(QUERY == gene_cluster))

#prepare the table with matadata for plot
meta_psbA_clust <- psbA_clust %>% 
  pivot_longer(!c("QUERY", "KO", "BRITE", "CAZY", "COG", "DISEASE", "DRUG", "ENZYME", "GO", "MODULE", "NETWORK", "PATHWAY", "PUBMED", "RCLASS", "REACTION", "TC", "VP", "DESCRIPTION", "length"), names_to = "sampleid", values_to = "counts", values_drop_na = 0) %>%
  merge(metadata)
meta_psbA_clust$Stromkilometer <- sub(",", ".", meta_psbA_clust$Stromkilometer)
meta_psbA_clust$Stromkilometer <- as.double(meta_psbA_clust$Stromkilometer)

#plot it!
meta_psbA_clust %>%  
  filter(Station != 'BunthausSpitze') %>% 
  filter(Sample_date != "Nov 21") %>%
  mutate(counts=counts/1000) %>%
  ggplot(aes(x=Stromkilometer, y=counts)) + ggtitle("psbA counts across different seasons and river sites") + geom_point() + 
  geom_tile() + facet_wrap(~data_type *Sample_date) + geom_smooth() 
#usually the highest variance in the beginning and at the end of the estuary 
#but very similar across  the samples
```
## 2.2 Check different fractions 
```{r}
#check different fractions
#plot it!
meta_psbA_clust %>%  
  filter(Station != 'BunthausSpitze') %>% 
  filter(Sample_date != "Nov 21") %>%
  mutate(counts=counts/1000) %>%
  ggplot(aes(x=Stromkilometer, y=counts)) + geom_point() +
  geom_tile() + facet_wrap(~ Sample_type) + geom_smooth()

#higher variance in particle associated - interesting opposite to the rhodopsins 
```

```{r}
#explore correlations
meta_psbA_clust_df <-  meta_psbA_clust  %>%  
  filter(data_type == "METAT") %>%
  filter(Station != 'BunthausSpitze') %>% 
  filter(Sample_date != "Nov 21") %>% 
  select(numeric_col) 

meta_psbA_clust_df[ meta_psbA_clust_df == ""] <- NA 

for (i in c(3:length(meta_psbA_clust_df[1,]))){
  meta_psbA_clust_df[,i] <-  sub(",", ".", meta_psbA_clust_df[,i]) #replace , to . to get a string which can be converted to double 
}

head(meta_psbA_clust_df)
#to numeric 
meta_psbA_clust_matrix <- matrix(apply(meta_psbA_clust_df,2, as.numeric), ncol=ncol(meta_psbA_clust_df), dimnames=list(c(),colnames(meta_rhd_clust_df)))
meta_psbA_clust_matrix[ is.na(meta_psbA_clust_matrix)] <- 0 
res <- cor(meta_psbA_clust_matrix)
res[ is.na(res)] <- 0 
round(res, 2)

heatmap(round(res, 2))
#no interresting correlations for PsbA counts
```
## 2.3 Variance analysis 
```{r}
#variance analysis 
var_meta_psbA_clust <- meta_psbA_clust  %>%  
  filter(Station != 'BunthausSpitze') %>% 
  filter(Sample_date != "Nov 21") %>% 
  group_by(Station, Sample_date) %>% 
  summarise(var = var(counts))

var_meta_psbA_clust_station <- meta_psbA_clust %>%  
  filter(Station != 'BunthausSpitze') %>% 
  filter(Sample_date != "Nov 21") %>% 
  group_by(Station) %>% 
  summarise(var = var(counts))

#variance across stations (similarly the variance is lower n the brakish water)
var_meta_psbA_clust_station  %>% arrange(factor(Station, levels = c("Muehlenberger Loch","Twielenfleth", "Schwarztonnensand", "Brunsbuettel","Meedem Grund"))) #change the order 
```

The variance is increasing with a station order 
